---
title: "Analysis 1: UNC Salaries"
author: "FIRSTNAME LASTNAME"
date:  "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

# Instructions

**Overview:** For each question, show your R code that you used to answer each question in the provided chunks. When a written response is required, be sure to answer the entire question in complete sentences outside the code chunks. When figures are required, be sure to follow all requirements to receive full credit. Point values are assigned for every part of this analysis. **Do not work with other students on this assignment. You are to complete this assignment by yourself.**

**Helpful:** Make sure you knit the document as you go through the assignment. Check all your results in the created HTML file.

**Submission:** Submit via an electronic document on Sakai. Must be submitted as an HTML file generated in RStudio. 

# Introduction

Universities are typically opaque, bureaucratic institutions. To be transparent to tax payers, many public schools, such as the University of North Carolina, openly report **[salary information](http://www.newsobserver.com/news/databases/public-salaries/)**. In this assignment, we will analyze this information from 2022 to answer pivotal questions that have endured over the course of time. The most recent salary data for UNC-Chapel Hill faculty and staff has already been downloaded in CSV format and titled *"Salary Data Export.csv"*.

To answer all the questions, you will need the R package `tidyverse` to make figures and utilize `dplyr` functions.


```{r setup, include=F}
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Data Information

Make sure the CSV data file is contained in the folder of your RMarkdown file. First, we start by using the `read_csv` function from the `readr` package found within the tidyverse. The code below executes this process by creating a tibble in your R environment named "salary".
```{r, message=F}
salary=read_csv("Salary Data Export.csv")
```

Now, we will explore the information that is contained in this dataset. The code below provides the names of the variables contained in the dataset.

```{r}
names(salary)
```

Next, we will examine the type of data contains in these different variables. 
```{r}
str(salary,give.attr=F)
```
You will notice that the variable "Hire_Date" is recorded as a character. The following code will first modify the original dataset to change this to a date variable with the format *mm/dd/yyyy*. Then, we will remove the hyphens to create a numeric variable as *yyyymmdd*. Finally, in the spirit of tidyverse, we will convert this data frame to a tibble.

```{r}
salary$Hire_Date=as.Date(salary$Hire_Date, format="%m/%d/%Y")
salary$Hire_Date=as.numeric(gsub("-","",salary$Hire_Date))
salary=as_tibble(salary)
```

Now, we will use `head()` to view of first five rows and the modifications made to the original data. The rest of the assignment will extend off this modified dataset named `salary` which by now should be in your global environment.
```{r}
head(salary,5)
```

# Assignment


## Part 1: Reducing the Data to a Smaller Set of Interest


### Q1 *(2 Points)* 

Create a new dataset named `salary2` that only contains the following variables:

- "Last_Name"

- "First_Name"

- "Age"

- "Hire_Date"

- "Category"

- "Salary"

- "Department"

- "Title"

Then, use the `names()` function to display the variable names of `salary2`.
```{r}
#
```

### Q2 *(2 Points)*

Now, we modify `salary2`. Rename the variables "Last_Name","First_Name", and "Hire_Date" to "Last","First", and "Hire", respectively. Do this for a new dataset called `salary3` and use `names()` to display the variable names of `salary3`.
```{r}
#
```

### Q3 *(2 Points)*

Now, we modify `salary3`. Create a new variable called "HireYear" that only contains the first four digits of the variable "Hire_Date" in a new dataset named `salary4`. Use the function `str()` to ensure that your new variable "HireYear" reports the year of the date that the employee was hired.

```{r}
#
```

### Q4 *(2 points)*

Now, we modify `salary4`. Create a new variable called "YrsEmployed" which reports the number of full years the employee has worked at UNC. Use the current year and assume that all employees are hired January 1. Create a new dataset named `salary5` and again use `str()` to display the variables in `salary5`.

```{r}
#
```

### Q5 *(4 points)*

Now, we modify `salary5` to create our final dataset named `salary.final`. Use the pipe `%>%` to make the following changes:

- Drop the variables "Hire", "HireYear", and "First". 

- Sort the observations first by "YrsEmployed" and then by "Salary", both in descending order from largest to smallest.

- Rearrange the variables so that "YrsEmployed" and "Salary" are the first two variables in the dataset, in that order, without removing any of the other variables.

After you have used the `%>%` to make these changes, use the function `head()` to display the first 12 rows of `salary.final`.

```{r}
#
```

## Part 2: Answering Questions Based on All Data

In these questions,find the answer using code, and then answer the question using complete sentences below the code.

### Q6 *(2 Points)*

What is the average salary of employees in my department (Statistics and Operations Research)?

Code *(1 Point)*:
```{r}
#
```

Answer *(1 Point)*: (Place Answer Here in Complete Sentences)

### Q7 *(4 Points)* 

How many employees have worked in History between 0 and 5 years (inclusive)?

Code *(2 Points)*:
```{r}
#
```

Answer *(2 Points)*: (Place Answer Here in Complete Sentences)

### Q8 *(4 Points)*

What is the difference between the average age of employees in the History department who are professors, associate professors, or assistant professors and the average age of employees in my department who are professors, associate professors, or assistant professors?

Code *(2 Points)*:
```{r}
#
```

Answer *(2 Points)*: (Place Answer Here in Complete Sentences)

## Part 3: Answering Questions Based on Summarized Data

### Q9 *(4 Points)*

Based off the data in `salary.final`, create a grouped summary based off combinations of "Department" and "Title". Call the new tibble `dept_title_summary`. Your summarized tibble, `dept_title_summary`, should report all of the following statistics with corresponding variable names in the following order.

- "n" = number of employees for each combination

- "salary" = average salary for each combination

- "age" = average age for each combination

- "corr" = correlation between the salary and age for each combination

In the process, make sure you use `ungroup()` with the pipe `%>%` to release the grouping so future work is no longer group specific. Also, the `cor()` can be used to calculate the correlation between two variables.

Following the creation of `dept_title_summary`, prove that your code worked by using `head()` to view the first 12 rows.

```{r,warning=FALSE}
#
```

### Q10 *(4 Points)*

Using the summarized data in `dept_title_summary`, use the `dplyr` functions to identify the 5 departments that award the highest average salary for assistant professors. **Only consider departments that have at least 3 assistant professors in their department.** The output should only show the names of the 5 departments along with the corresponding average salary.

Code:
```{r}
#
```

### Q11 *(4 points)*

Create a scatter plot using `geom_point()` along with fitted lines using `geom_smooth` with the argument `method="lm"` showing the linear relationship between average salary and the average age. Put the average age on the x-axis of your plot. **For this plot, use the summarized data in `dept_title_summary` and exclude the combinations of Department and Title where there are less than 5 employees.**

Following the plot, please explain what this plot suggests about the relationship between the average salary and the average age. Make reference to the figure and use descriptive adjectives (i.e. "strong", "weak", etc.) and terms (i.e. "positive", "negative", etc.) that are appropriate for discussing linear relationships.

Code and Figure *(2 Points)*:
```{r}
#
```

Answer *(2 Points)*: (Place Answer Here in Complete Sentences)

### Q12 *(6 Points)*

The purpose of summarizing the data was to analyze the previously discussed linear relationship by group. In `dept_title_summary`, there are 750 unique departments represented. You can verify this by using `length(unique(dept_title_summary$Department))`. In this part, I want you to select 3 academic departments, not previously discussed, and in one figure, display the scatter plots and fitted linear regression lines representing the relationship between average salary and average age in 3 different colors. Remove the shading of the confidence intervals for the 3 regression lines. Then, in complete sentences, I want you to state what departments you chose and explain the differences and/or similarities between the groups regarding the previously mentioned relationship. 


Code and Figure: *(4 Points)*:
```{r}
#
```

Answer *(2 Points)*: (Place Answer Here in Complete Sentences)
