---
title: "Lab 6"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  4 (Pg. 302); 1 (Pgs. 316-317); 1 (Pgs. 328-329); 1, 2 (Pgs. 353-354)

**Submission:** Submit via an electronic document on Sakai. Must be submitted as a HTML file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 16 (Pg. 302)

##  Exercise 4

### a)
```{r}
#
```

### b)
```{r}
#
```

### c)
```{r}
#
```

### d)
```{r}
#
```

# Chapter 17 (Pgs. 316-317)

##  Exercise 1

### a)
```{r}
#
```

### b)
```{r}
#
```

### c)
```{r}
#
```

### d)
```{r}
#
```

# Chapter 17 (Pgs. 328-329)

##  Exercise 1

### a)
```{r}
#
```

### b)
```{r}
#
```

### c)
```{r}
#
```

### d)
```{r}
#
```

# Chapter 18 (Pgs. 353-354)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```
