---
title: "STOR 320 Tutorial on Modeling 2"
author: "Mario Giacomazzo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)    #Essential Functions
library(modelr)       #Helpful Functions in Modeling
DATA=read_csv("AirWaterTemp.csv",col_types=cols()) #River Data
```

# Introduction

Today, we will work with daily water temperature and air temperature data observed for `r length(unique(DATA$L))` rivers in Spain. The goal of this tutorial is to identify the best model for predicting the maximum water temperature given the maximum air temperature. In the preview below, `W` represents the daily maximum water temperature and `A` represents the daily maximum air temperature. The data contains almost a full year of data for each of the `r length(unique(DATA$L))` different rivers.

```{r,echo=F}
head(DATA)
```


#Part 1: Examining the Relationship

##Chunk 1: Overall Relationship
```{r,echo=F,eval=T,message=F}
ggplot(data=DATA) +
  geom_point(aes(x=A,y=W),alpha=0.3)+
  geom_smooth(aes(x=A,y=W)) +
  theme_minimal()
```


##Chunk 2: Location-Specific Relationship

```{r,echo=F,eval=F}
Location = 103

DATA %>% filter(L==Location) %>%
  ggplot()+
    geom_point(aes(x=A,y=W),alpha=0.3)+
    geom_smooth(aes(x=A,y=W)) +
    theme_minimal()
  
```

##Chunk 3: Split Data into Train and Test Sets
```{r,eval=F}
set.seed(INTEGER)
TEST.LOCATIONS=sample(x=unique(DATA$L),size=3,replace=F)

TRAIN = anti_join(DATA,tibble(L=TEST.LOCATIONS),by="L")
TEST = semi_join(DATA,tibble(L=TEST.LOCATIONS),by="L")
```

##Chunk 4: Plots of Relationship for Train and Test Data
```{r,echo=F,eval=F}
WAPLOT2.func=function(DATA){
  ggplot(data=DATA)+
    geom_point(aes(x=A,y=W),alpha=0.3)+
    geom_smooth(aes(x=A,y=W)) +
    theme_minimal()
}

WAPLOT2.func(TRAIN)
WAPLOT2.func(TEST)
```


#Part 2: Linear Regression Model

##Chunk 1: Fitting Linear Model to Train Data
```{r,echo=F,eval=F}
linmod=lm(W~A,data=TRAIN)
summary(linmod)
```

##Chunk 2: Getting Predictions from Linear Model
```{r,eval=F}
TRAIN2 = TRAIN %>% add_predictions(linmod,var="linpred")
TEST2 = TEST %>% add_predictions(linmod,var="linpred")
```

##Chunk 3: Getting Residuals from Linear Model
```{r,eval=F}
TRAIN3 = TRAIN2 %>% add_residuals(linmod,var="linres")
TEST3 = TEST2 %>% add_residuals(linmod,var="linres")
```

#Part 3: Polynomial Regression Model

##Chunk 1: Fitting Polynomial Regression Models
```{r,eval=F}
poly2mod=lm(W~A+I(A^2),data=TRAIN)
poly3mod=lm(W~A+I(A^2)+I(A^3),data=TRAIN)
poly4mod=lm(W~A+I(A^2)+I(A^3)+I(A^4),data=TRAIN)
anova(linmod,poly2mod,poly3mod,poly4mod,test="Chisq")
```

##Chunk 2: Getting Predictions from Polynomial Models
```{r,eval=F}
TRAIN4 =TRAIN3 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
  
TEST4 =TEST3 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
```

##Chunk 3: Getting Residuals from Polynomial Models
```{r,eval=F}
TRAIN5 =TRAIN4 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
  
TEST5 =TEST4 %>% 
  add_predictions(poly2mod,var="poly2pred") %>%
  add_predictions(poly3mod,var="poly3pred") %>%
  add_predictions(poly4mod,var="poly4pred")
```
