---
title: "STOR 320 Tutorial on Programming 1"
author: "Mario Giacomazzo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)    #Essential Functions
library(Ecdat)        #Contains Economic Datasets
```

# Introduction

In this tutorial, we will try to understand some fundamental control structures used in statistical programming. In the beginning, we will separately analyze different control structures. The ultimate goal is to combine these methods to solve all our computing and life problems. If you can imagine it, R can do it. Programming R is downright magical. If Morpheus calls, pick up the phone.

# Part 1: *If-Else* Statements

## Chunk 1: Illustration of *If*
```{r,eval=F}
x = 3
if(x > 0){
  print(log(x))
}

x = -3
if(x > 0){
  print(log(x))
}
```

## Chunk 2: Illustration of *If-Else*
```{r,eval=F}
x = 3
if(x > 0){
  print(log(x))
} else{
  message("Unable to Take Logarithm")
}

x = -3
if(x > 0){
  print(log(x))
} else {
  message("Unable to Take Logarithm")
}
```

## Chunk 3: Potential Problem of *If-Else* Statements
```{r,eval=F}
x = BLANK
if(x > 0){
  print(log(x))
}

if(x > 0){
  print(log(x))
} else{
  message("Unable to Take Logarithm")
}
```

##Chunk 4: Fixing Potential Problem in Chunk 3
```{r,eval=F}
x=BLANK
if(is.numeric(x)){
  if(x > 0){
    print(log(x))
  } else{
    message("Unable to Take Logarithm")
  }
} else{
  message("No Strings Attached")
}
```

##Chunk 5: Vectorized Version with `ifelse()`
```{r,eval=F}
x=c(-1,3,200)
print(log(x))

y1 =  if(x > 0){
        log(x)
      } else{
        NA
      }
print(y1)

y2 = ifelse(x>0,log(x),NA)
print(y2)
```

##Chunk 6: Nested `ifelse()` Statements
```{r,eval=F}
x=rnorm(1000,mean=0,sd=1)
y=ifelse(abs(x)<1,"Within 1 SD",ifelse(abs(x)>2,"Far Far Away","Between 1 and 2 SD"))
y.fct=factor(y,levels=c("Within 1 SD","Between 1 and 2 SD","Far Far Away"))
ggplot() +
  geom_bar(aes(x=y.fct),fill="lightskyblue1") +
  theme_minimal()
```

#Part 2: Loops

##Chunk 1: Checking Geometric Series Proof with *for* loop
```{r,eval = F}
a=BLANK #Any Number
r=BLANK #Any Number Between -1 and 1: abs(r)<1

theoretical.limit=a/(1-r)

START=a

FINISH.1 = START + a*r^1

FINISH.2 = FINISH.1 + a*r^2

FINISH.3 = FINISH.2 + a*r^3

FINISH.10 = a 
for(k in 1:10){
  FINISH.10=FINISH.10+a*r^k
}

FINISH.100 = a 
for(k in 1:100){
  FINISH.100=FINISH.100+a*r^k
}

DATA = tibble(k=c(1,2,3,10,100,"Infinity"),
            SUMMATION=c(FINISH.1,FINISH.2,FINISH.3,
                        FINISH.10,FINISH.100,
                        theoretical.limit))
print(DATA)

ABSOLUTE.ERROR = abs(FINISH.100-theoretical.limit)
print(ABSOLUTE.ERROR)
```

##Chunk 2: Checking Geometric Series Proof with *while* loop
```{r,eval=F}
a=BLANK
r=BLANK

FINISH=a
k=0
while(abs(FINISH-a/(1-r)) > 1e-10) {
  k=k+1
  FINISH = FINISH + a*r^k
  #if(k>100) break
}
print(c(k,FINISH))
```

##Chunk 3: Saving Steps in Geometric Series for Figure
```{r,eval=F}
a=10
r=-0.75
theoretical.limit=a/(1-r)

K=10 #How Many Steps Do You Want to Save?

summation=rep(NA,(K+1))
summation[1]=a
for (k in 1:K) {
  summation[k+1]=summation[k] + a*r^k
}

ggplot() +
  geom_line(aes(x=1:(K+1),y=summation)) +
  geom_hline(yintercept=theoretical.limit,
             linetype="dashed")
```
