---
title: "Lab 7: Relational Data"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
# This is a good place to put libraries required for using the ggplot function
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(tidyverse)    #Loads the tidyverse suite of packages
#install.packages('Lahman')
library(Lahman)       #Contains the baseball data to use
library(nycflights13) #Contains the 2013 NYC flights data
#install.packages("maps")
library(maps)
```

# Introduction

The main purpose of this lab is to practice data join skills from Chapter 10. The functions and their purposes are listed as follows:

- `inner_join()` Keeps observations appear in both datasets.

- `left_join()` Keeps all observations in left dataset.

- `right_join()` Keeps all observations in right dataset.

- `full_join()` Keeps all observations in both datasets.

- `semi_join()` Keeps all observations in left dataset that have a match in right dataset.

- `anti_join()1` Drops all observations in left dataset that have a match in right dataset.

You will need to modify the code chunks so that the code works within each of chunk (usually this means modifying anything in ALL CAPS). You will also need to modify the code outside the code chunk. When you get the desired result for each step, change `Eval=F` to `Eval=T` and knit the document to HTML to make sure it works. After you complete the lab, you should submit your HTML file of what you have completed to Canvas before the deadline.

# Exercises

## Part 1

In part 1, you will practice the skills using the datasets from the R package `Lahman`. This database includes data related to baseball teams. It includes summary statistics about how the players performed on offense and defense for several years. It also includes personal information about the players.

The `Batting` data frame contains the offensive statistics for all players for many years. You can see, for example, the top 10 hitters in 2016 by running this code: (For more details of the dataset run `?Batting` in console.)
```{r}
top <- Batting %>% 
  filter(yearID == 2016) %>%
  arrange(desc(HR)) %>%
  slice(1:10)

top #Do Not Remove
```

But who are these players? We see an ID, but not the names. The player names are in this table named `People`.

```{r}
head(People,5)
```

We can see column names `nameFirst` and `nameLast` in table `People`.

### 1.1 (1 Point)

Use the `left_join` function to create a data frame called `top1`, which contains information of the 10 top home run hitters. The table should have the following columns: `playerID`, `nameFirst`, `nameLast`, and number of home runs (`HR`).

```{r, eval=F}
top1 = top %>% 
  left_join(DATA, by = KEY) %>%
  select(VARIABLES)

top1 #Do Not Remove
```

### 1.2 (2 Points)

Data `Salaries` contains the baseball player salary data. 

```{r}
head(Salaries,5)
```

You may be curious about the salaries of the top 10 hitters in 2016:

- Now create a new data frame called `top2` by adding top 10 hitters' salaries to `top1` and including only `nameFirst`, `nameLast`, `teamID`, `HR`, and `salary` columns. 
- Rename the columns to `FirstName`, `LastName`, `Team`, `Homeruns` and `Salary` respectively.
- Arrange the data frame by `Salary` in descending order. 

Note that salaries are different every year so make sure to filter for the year 2016. This time, only use `right_join` to complete the exercise.

```{r, eval=F}
top2 = DATA %>%
  filter(CONDITION) %>%
  right_join(DATA,by=KEY) %>%
  select(SELECT_AND_RENAME) %>%
  arrange(desc(VAR))

top2   #Do Not Remove
```

## Part 2

In this part, we will explore relational data from `nycflights13`, which contains four data frames related to the `flights` table that you used in previous assignments.

### 2.1 (2 Points)

Data `airports` gives information about each airport, such as latitude and longitude, identified by the `faa` airport code.
```{r}
head(airports,5)
```

Based on `flights`, compute the average arrival delay by destination (`dest`) and ignore missing values, then join on the `airports` data frame. Only use an inner join.

```{r, eval=F}
delay = DATA %>%
  group_by(VAR) %>%
  summarise(avg_arr_delay=COMPUTATION,.groups='drop') %>%
  inner_join(DATA,by=KEY) %>%
  select(VARS)
delay #Do Not Remove
```

### 2.2 (1 Points)

Draw a scatterplot with dots representing destination locations and colors of dots representing average arrival delay on US map. We do this to look at the spatial distribution of average arrival delay.
```{r, eval=F}
delay %>%
  ggplot(aes(SPECIFY_AESTHETICS)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

### 2.3 (1 Points)
Data `planes` gives information about each plane, identified by its `tailnum`. Note that `year` column in `planes` represents the year a plane was manufactured, which is different from `year` column in `flights`.

```{r}
head(planes,5)
```

Use the `planes` data to calculate the `age` of planes, assuming current year is 2013. Keep only `tailnum` and `age` in the output table `plane_ages`. 

```{r, eval=F}
plane_ages <- 
  DATA %>%
  mutate(age = FUNCTION) %>%
  select(VARS)
plane_ages #Do Not Change
```

### 2.4 (2 Points)

Is there a relationship between the age of a plane and its delays? 

- Join the `plane_ages` with `flights`, keeping observations with matches in both datasets.
- Summarize the average departure delay by plane `age` and ignore missing values.
- Draw a scatterplot of plane age vs. average departure delay.

```{r, eval=F}
DATA %>%
  inner_join(DATA, by = KEY) %>%
  group_by(VAR) %>%
  summarise(avg_dep_delay = FUNCTION, .groups='drop') %>%
  ggplot(aes(SPECIFY_AESTHETICS)) +
  geom_point()
```


### 2.5 (0.5 Points)

What does it mean for a flight to have a missing `tailnum`? 
```{r, eval=F}
flights %>%
  filter(is.na(VAR))
```
Answer: ANSWER_HERE

### 2.6 (0.5 Points)

What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems. Check the documentation of `planes` for help.) 

```{r, eval=F}
DATA %>% 
  anti_join(DATA,by=KEY) %>%
  count(VAR) %>%
  arrange(desc(n))
```

Answer: ANSWER_HERE