---
title: "Lab 10: Modeling Basics"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)
library(modelr)
library(broom)
```

# Introduction

In this lab, you will build predictive models for board game ratings. The dataset below was scraped from [boardgamegeek.com](www.boardgamegeek.com) and contains information on the top 4,999 board games. Below, you will see a preview of the data

```{r}
bgg<-read.csv("bgg.csv")
bgg2=bgg[,c(4:13,15:20)]
head(bgg2)
```


# Board Game Analysis

### Q1 (1.5 Points)

There are 16 variables and we want to create some more. Create a new dataframe called $bgg3$ where you use the mutate function to create the following variables:

- *duration=2018-year+1*
- *vote.per.year=num_votes/duration*
- *own.per.year=owned/duration*
- *player.range=max_players-min_players*
- *log_vote=log(num_votes+1)*
- *log_own=log(owned+1)*
- *diff_rating=avg_rating-geek_rating*

```{r,eval=F}

head(bgg3)
```

**Question:** In complete sentences, what is the purpose of adding 1 for the log transformed variables?

YOUR ANSWER IN COMPLETE SENTENCES

**Question:** In complete sentences, what is the purpose of adding 1 in the creation of the year variable?

YOUR ANSWER IN COMPLETE SENTENCES

### Q2 (2 Points)

We hypothesize the geek rating increases when the number of votes increases and/or the ownership increases. Create four scatter plots showing the association with geek_rating and the following variables:

- *num_votes*
- *owned*
- *log_vote*
- *log_own*


```{r,eval=F}

```

**Question:** In complete sentences, describe how the relationship changes when you take the log of the independent variable.

YOUR ANSWER IN COMPLETE SENTENCES



### Q3 (0.5 Points)

Randomly sample approximately 80\% of the data in `bgg3` for a training dataset and the remaining will act as a test set. Call the training dataset `train.bgg` and the testing dataset `test.bgg`.

```{r,eval=F}
set.seed(COMPLETE)

bgg4= bgg3 %>%
        mutate(Set=sample(COMPLETE))

train.bgg<-filter(bgg4,Set=="Train")
test.bgg<-filter(bgg4,Set=="Test")

```



### Q4 (0.5 Points)

Now, we want to fit models to the training dataset. Use the `lm()` function to create 3 model objects in R called `lm1`, `lm2`, `lm3` based on the following linear models, respectively:

- $\textrm{geek_rating}=\beta_0+\beta_1 log(\textrm{num_votes})+\epsilon$
- $\textrm{geek_rating}=\beta_0+\beta_1 log(\textrm{owned})+\epsilon$
- $\textrm{geek_rating}=\beta_0+\beta_1 log(\textrm{owned})+ \beta_2 \textrm{vote.per.year}+ \beta_3 \textrm{weight} + \epsilon$

```{r,eval=F}
lm1 = lm(COMPLETE,data=train.bgg)
lm2 = lm(COMPLETE,data=train.bgg)
lm3 = lm(COMPLETE,data=train.bgg)
```

### Q5 (1 Point)

Add predictions and residuals for all 3 models to the test set. Create a new data frame called `test.bgg2` and give all your predictions and residuals different names. Use the `str()` function to show these variables were created


```{r,eval=F}


str(test.bgg2)
```


### Q6 (0.5 Points)

Create a function called `MAE.func()` that returns the mean absolute error based on a vector of the residuals and test your function on the vector called `test`.

Solution 1:
```{r,eval=F}
test=c(-5,-2,0,3,5)



MAE.func(test)
```

### Q7 (1 Point)

Use your function on the `test.bgg2` to calculate the out-of-sample MAE of all three models based on the associated residuals. Make sure you display the mean absolute error from these different models in your output.

```{r,eval=F}

```

**Question:** Which model does the best job at predicting the geek rating of these board games?

YOUR ANSWER IN COMPLETE SENTENCES

### Q8 (3 Points)

For the third model only, use 10-fold cross-validation and measure the out-of-sample mean absolute error. Print out the final cross-validated mean absolute error.

```{r,eval=F}


```

**Question:** What is the absolute difference between the out-of-sample mean absolute error measured using a test set and the mean absolute error measured using cross validation? When you type your answer in complete sentences use inline R code to calculate the absolute difference and input it directly into your sentence.

YOUR ANSWER IN COMPLETE SENTENCES





