---
title: "Lab 11: Modeling for Final Project"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)
```

# Introduction

In this lab, you will work on the final paper with your groups. Every member of the group should complete one of these documents and submit this to Canvas. The work from each group member should be different, and you should work together to ensure that the work is different. The code chunk below is for loading the data into R and for potentially cleaning the data. 

```{r}
#
```


## Q1 (4 Points)

Based on the requirements of the final paper, your group should have at least one variable that your group believes would be useful to predict. This could be a categorical variable or a numeric variable. I recommend focusing on a numeric variable for this lab unless your group only wants to build classification models for your final paper. 

In the code chunk below, I want each member to create a professional visual regarding this variable of interest (the variable you want to predict). This should be a visual that involves at least 2 other variables in the dataset. This should be a visual that could potential be used in the "Data" section of the final paper to introduce this variable of interest to your audience and to build a case for predicting this variable based off other information in the dataset. This visual can also be used to highlight a significant problem in the data pertaining to this variable that your audience needs to know about prior to building models and making prediction.

**Remember: Each member of the group should have a different picture. You are helping each other, but working independently.**

```{r}
#
```

After each member creates a different visual, I want you all to discuss your visuals with each other. When working in a group, it is important that everyone actively tries to contribute helpful things to the project, but this doesn't mean that your work will be included in a final paper. After defending your visuals to each other, which group member do you believe made a visual that has the best chance of being included in the final paper? **I don't want you to pick yourself**

**Who made the most useful visual according to you?** REPLACE ALL CAPS WITH THE NAME OF THE PERSON

## Q2 (2 Points)

In this code chunk below, I want each member to make one model designed to predict the variable of interest. (If your group picked a categorical/binary variable, then this question is really about classification.) You should tell your group what predictor variables are going to be in your model and what type of model you are going to fit to ensure that every member of the group has a different model. The models could be different based off the subset of predictor variables (different sets of X variables) or based off the type of model (linear regression, random forest, polynomial, generalized additive models, etc.). Try to use interactions between different variables and try to engineer new variables that could potentially be helpful in predicting the variable of interest. 

I want each member of the group to pick a different model, but each group member should be fitting a model that they think will potentially do a good job at prediction. 

In the code chunk below, I want you to start by fitting the model you chose to the whole dataset. Use a function like `summary()` to show the output of the fitted model. If you have to do data cleaning or other coding to get the dataset ready for the model you chose, you can include this in this code chunk; however, I only want to see output pertaining to the model you fit. 

```{r}
#
```

## Q3 (4 Points)

Perform Leave-One-Out Cross-Validation to estimate the out-of-sample RMSE of the model you chose in the previous question.

1. Start by creating a variable in the data called "Predict" that is all missing/NA.
2. Create a loop to loop through each row of the dataset.
3. For each iteration of the loop, you should remove the corresponding row from the data. 
4. Fit the model to the remaining data. 
5. Use the predict function with the fitted model to make a prediction for the row that was ignored.
6. Replace the NA in the "Predict" column of the dataset corresponding with the row that was removed prior to fitting the model. 

```{r}
#
```

Help each other with this part!! Once you figure out  as a group how you are going to code this, you can all use similar code on different models. After you all complete this part, compare the RMSE of all the models that your group came up with and identify the group member who made the "best" model. Remember the "best" model may not be the model that simply has the lowest cross-validated RMSE. Try to seek a model that is simple, adequately predicts, and could be useful to your audience who may implement this. You all should discuss your results with each other to identify the group member who the group believes made the "best" model. 

**Who made the model that the group identified as "best"?** REPLACE ALL CAPS WITH THE NAME OF THE PERSON

## Q4 (0 Points)

After working with your team on this lab, are there any clear problems with any of the group members. Was there a group member who didn't contribute? Was there a group member who illustrated an inability to code on their own? Was there a group member who made no effort in trying to create a good visual or good model? Was there a group member who disregarded the ideas of others in favor of their own? Was there a group member who refused to participate? 

If there is a group member(s), who is clearly a problem, what would be the idea way to handle it? Do you think this person should remain in the group based off their contribution from the beginning of the semester? Should I have this person write their own final paper? Should I email or talk to this student? 

REPLACE ALL CAPS WITH YOUR THOUGHTS

**If you provided an answer to this question, email me and consider having a meeting with me on zoom to maintain privacy. The only way I will react to what you write here is if we have a meeting to discuss this.**