---
title: "Lab 8: More Relational Data"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---


```{r setup, include=F}
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
```

# Linking Important Information to Violent Crimes Data

In this lab, we will join all 5 datasets from the lecture series on web scraping. In the zipped folder, there are *5* CSV files. In this section, we are going to merge all of that data into one object called `FINAL.VIOLENT`. 

```{r,message=F}
# Data for Part 1
VIOLENT=read_csv("FINAL_VIOLENT.csv")
ZIP=read_csv("FINAL_ZIP.csv")
STATE_ABBREV=read_csv("FINAL_STATE_ABBREV.csv")
CENSUS=read_csv("FINAL_CENSUS.csv")
S_VS_D=read_csv("FINAL_SAFE_VS_DANGEROUS.CSV")
```

## 1: Get Desirable Data (1 Point)

The dataset `S_VS_D` contains a variable `CLASS` where "S=Safe" and "D=Dangerous" according to the article *[These Are the 2018 Safest and Most Dangerous States in the U.S](https://www.securitysales.com/fire-intrusion/2018-safest-most-dangerous-states-us/)* by [Steve Karantzoulidis](https://www.securitysales.com/author/stevek/). We seek to compare the violent crime statistics for states not in this list. Use a filtering join to create a new data frame called `VIOLENT2` that only contains violent crime statistics from the states not represented in the data frame `S_VS_D`.  Use `str(VIOLENT2)` to display the variables and the dimensions of `VIOLENT2`. Do this without renaming any of the variables!!


```{r, eval=F}
VIOLENT2 = anti_join(COMPLETE,by=COMPLETE)

str(VIOLENT2) #Do Not Change
```

## 2: Fixing Instances in Crime (1 Point)

Start by creating a new data set called `VIOLENT3` based on `VIOLENT2` that fixes some problems in the variable `City`. Specifically, we would like to change "Louisville Metro" and "Nashville Metropolitan" to "Louisville" and "Nashville", respectively. 

```{r, eval=F}
VIOLENT3=VIOLENT2 %>% mutate(City=ifelse(COMPLETE),
                             City=ifelse(COMPLETE))

str(filter(VIOLENT3,City %in% c("Louisville","Nashville"))) #Do Not Change
```

## 3: Merging Data (2 Points)

Next, create a new data frame named `VIOLENT4` that merges the population change and density measures from 2019 contained in `CENSUS` into `VIOLENT3` based on city and state. Use `head(VIOLENT4)` to give a preview of the new merged dataset. Only use a `left_join` for this part.

```{r, eval=F}
VIOLENT4 = left_join(COMPLETE)

head(VIOLENT4) #Do Not Change
```

### 3.1: Why is the Washington D.C. missing CENSUS information? Answer in complete sentences. (1 Point)

PLACE YOUR ANSWER HERE

## 4: Merging more Data (2 Points)

Either ambitiously using one-step or less ambitiously using multiple steps merge the longitude and latitude information provided in `ZIP` into  `VIOLENT4` based on city and state. You will need to use `STATE_ABBREV` data to link these two data frames. Your final data frame named `FINAL.VIOLENT` should contain all of the information in `VIOLENT4` along with the variables `lat` and `lon` from `ZIP`. There should be **no** state abbreviations in `FINAL.VIOLENT` since this information is redundant.  Only use `left_join` and do this without renaming a single variable. Use `str(FINAL.VIOLENT)` to preview table.

```{r, eval=F}
FINAL.VIOLENT = COMPLETE

str(FINAL.VIOLENT) #Do Not Change
```

### 4.1: Why is the Washington D.C. missing latitude and longitude information? Answer in complete sentences. (1 Point)

PLACE YOUR ANSWER HERE

## 5: Fixing Washington DC (2 Points)

It is not possible to fix Washington DC, but lets try anyway.

In `FINAL.VIOLENT`, we don't have census information or location information for Washington DC. Make modifications to the datasets, `CENSUS` and `STATE_ABBREV`, so that when you merge the datasets again, Washington DC has both census and location information. Then, redo the merges and create a new data frame called `FINAL.VIOLENT.FIX` that is not missing any information. You will probably have to do this in multiple steps. Only use `left_join` for all merges and never rename any variables.

Finally, print out the row of the data pertaining to Washington DC, so that your lab instructor can see that you did it correctly. 

```{r, eval=F}
COMPLETE


filter(FINAL.VIOLENT.FIX,City=="Washington",State=="District Of Columbia") # Do Not Change
```