---
title: "Summer Lab 4"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
# This is a good place to put libraries required for using the ggplot function
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)
library(Ecdat)
library(modelr)
library(broom)
```

# Lab 4a: Control Structures

## Introduction

The main purpose of this lab is to practice control structures in R:

- `if` and `else`: testing a condition and acting on it
- `for`: execute a loop a fixed number of times
- `while`: execute a loop while a condition is true
- `repeat`: execute an infinite loop (must break out of it to stop) • break: break the execution of a loop
- `next`: skip an iteration of a loop

You will need to modify the code chunks so that the code works within each of chunk (usually this means modifying anything in ALL CAPS). You will also need to modify the code outside the code chunk. When you get the desired result for each step, change `Eval=F` to `Eval=T` and knit the document to HTML to make sure it works. After you complete the lab, you should submit your HTML file of what you have completed to Canvas before the deadline.

## Part 1: Vector and Control Structures

### 1.1 (1 points) 

Write code that creates a vector `x` that contains `100` random observations from the standard normal distribution (this is the normal distribution with the mean equal to `0` and the variance equal to `1`). Print out only the first five random observations in this vector.

```{r}
#
```

### 1.2 (1 points) 

Write code that replaces the observations in the vector `x` that are greater than or equal to `0` with a string of characters `"non-negative"` and the observations that are smaller than `0` with a string of characters `"negative"`. Hint: try `ifelse()` funtion. Print out the first five values in this new version of `x`.

```{r}
#
```

### 1.3 (1 points) 

Write `for`-Loop to count how many observations in the vector `x` are non-negative and how many observations are negative. (There are many easier ways to solve this problem. Use `for`-Loop or get 0 points. Use the `cat()` function to print out a sentence that states how many non-negative and negative obervations there are. For example, "The number of non-negative observations is 32".

```{r}
#
```

## Part 2: Matrix and Control Structures

### 2.1 (2 points) 

Create a $100000$ by $10$ matrix `A` with the numbers $1:1000000$. The first row of this matrix should be the numbers 1 to 10. The second row of this matrix should be the numbers 11 to 20. Create a `for`-loop that calculates the sum for each row of the matrix and save the results to a vector `sum_row` and print out the first five values of `sum_row`.

```{r,eval=F}
A = matrix(1:1000000, COMPLETE) # DO NOT CHANGE
```

Verify that your results are consistent with what you obtain with the built-in `rowSums` function. 
```{r,eval=F}
sum_row_rowSums = as.integer(rowSums(A))
sum_row_rowSums[1:5]
```

### 2.2 (2 points) 

Another common loop structure that is used is the `while` loop, which functions much like a `for` loop, but will only run as long as a test condition is `TRUE`. Modify your `for` loop from the previous exercise and make it into a `while` loop. Use the `identical()` function to check if the results from the `for` loop are the same as the results from `while` loop.

```{r}
#
```

## Part 3: Data Frame and Control Structures

### 3.1 (2 points) 

Write a `for` loop to compute the mean of every column in `mtcars` and save the results to a vector `col_mean`. Ignore missing values when taking the mean.

```{r}
#
```

### 3.2 (1 points) 

Compute the number of unique values in each column of `iris` and print the results during a loop. Use the `cat()` function to print out the values in a sentence with the corresponding name of the variable. For example, "The number of unique values for Sepal.Length is 35".

```{r}
names(iris) #DO NOT CHANGE
```

# Lab 4b: Programming

## Introduction

In this lab, you will practice writing functions and using loop functions in R. The loop functions are:

- `lapply()`: Loop over a list and evaluate a function on each element

- `sapply()`: Same as `lapply` but try to simplify the result

- `apply()`: Apply a function over the margins of an array

- `tapply()`: Apply a function over subsets of a vector

- `mapply()`: Multivariate version of `lapply`

## Simulation study
Suppose that $X_1,\ldots,X_n$ are independent and identically distributed (iid) binomial random variables such that
\[
  P(X_i=x\mid k,p)
  ={k\choose x}p^x(1-p)^{k-x},\quad x=0,1,\ldots,k
\]
for all $i=1,\ldots,n$. Assume that both $k$ and $p$ are unknown and use the method of moments to obtain point estimators of both parameters. This somewhat unusual application of the binomial model has
been used to estimate crime rates for crimes that are known to have many unreported
occurrences. For such a crime, both the true reporting rate, $p$, and the total number
of occurrences, $k$, are unknown. Equating the first two sample moments to those of the population yields the system
of equations
\[
  \bar X=kp
  \quad\text{and}\quad
  \frac1n\sum_{i=1}^nX_i^2=kp(1-p)+k^2p^2,
\]
where $\bar X$ is the sample mean. Solving for $k$ and $p$ leads to
\[
  \hat k=\frac{\bar X^2}{\bar X-(1/n)\sum_{i=1}^n(X_i-\bar X)^2}
  \quad\text{and}\quad
  \hat p=\frac{\bar X}{\hat k}.
\]
It is difficult to analyze the performance of $\hat k$ and $\hat p$ analytically so you are asked to perform a simulation study using `R`. The idea is to generate random samples and investigate the performance of $\hat k$ and $\hat p$ using random samples.

### 1.1 (1 point)

Generate a single simple random sample vector `x` of length `n = 50` from the binomial distribution with the parameters `k = 10`, `p = 0.4`. 

```{r, eval=F}
k = VALUE
p = VALUE
x = rbinom(50,k,p)

hist(x) #Do Not Change
```

### 1.2 (4 points)

Write a function that takes a sample vector as its input and returns the estimates of `k` and `p` given above. Observe the output of your function and make sure your estimates of $k$ and $p$ are close to the truth.

```{r, eval=F}
est_kp = function(x){
  X_bar = COMPLETE
  n = length(x)
  k_hat = COMPLETE
  p_hat = COMPLETE
  return(c(k_hat,p_hat))
}

est_kp(rbinom(5000,k,p)) #Do Not Change
```

### 1.3 (4 points)

Generate `N = 1000` samples of size `n = 50` (as in the first question) and calculate `N = 1000` estimates of $k$ and `N = 1000` estimates of $p$. Please use Loop functions (apply, etc.) in this part. Make sure `rbinom` and `est_kp` functions are used in this part. Also, observe the output of the first ten samples. You will see the estimates of $k$ in the first row and the estimates of $p$ in the second row.

```{r, eval=F}
N = 1000
data = matrix(rbinom(N*50,k,p),N)
E50 = apply(COMPLETE,COMPLETE,COMPLETE)

E50[,1:10] #Do Not Change
```

### 1.4 (2 points)

Repeat Question 3 when `n <- 100` and when `n <- 250`. 

```{r, eval=F}
E100 = COMPLETE
E250 = COMPLETE

E100[,1:10] #Do Not Change
E250[,1:10] #Do Not Change
```

### 2.1 (3 points)

Estimate the bias and the mean squared error (MSE) of $\hat k$ and the bias and the MSE of $\hat p$ for each sample size (`n <- 50`, `n <- 100` and `n <- 250`). Do the estimators seem to overestimate or underestimate the parameters? Think about this and answer the follow-up question.

```{r, eval=F}
# bias
rowMeans(FILL)-c(10,.4)
COMPLETE_THE_OTHER_TWO

# mse
rowMeans((FILL-c(10,.4))^2)
COMPLETE_THE_OTHER_TWO
```

*Question:* How do the bias and the MSE change when the sample size increases?

REPLACE YOUR ANSWER HERE WITH COMPLETE SENTENCES

### 3.1 (3 points)

Make a single plot using `ggplot2` that contains three box plots of the estimates of the parameter $k$ when `n = 50`, `n = 100`, `n = 250` (the first from the left box plot has to describe the estimates when `n = 50`, the second from the left box plot has to describe the estimates when `n = 100` and the third from the left box plot has to describe the estimates `n = 250`). Include the true value of the parameter as a red horizontal line (`geom_hline()` and use the argument `color`) and label the plot appropriately.

You will need to construct a dataset to do this which is the point of the first part of the code. Run the code in parts to see what is happening.

```{r,eval=F}
df_k<-tibble(
  n=factor(rep(c("50","100","250"),each=N),c("50","100","250")),
  Estimates=c(COMPLETE)
  )
ggplot(data = df_k, mapping = aes(x = FILL, y = FILL)) +
	FUNCTION +
	geom_hline(yintercept = k, colour = "red")
```

### 3.2 (3 points)

The estimates $\hat k$ can result in values that are far away from the true value of the parameter when the sample size is small and the box plots might not be particularly informative in such a situation. Remove the estimates from the plot that are outside of the interval $[0,50]$ so that the box plots are more informative.

You are redoing the plot from the previous question without extreme estimates for $k$.

```{r, eval=F}
ggplot(
  data = filter(df_k, CONDITION), 
  mapping = aes(x = FILL, y = FILL)
  ) +
	FUNCTION +
	geom_hline(yintercept = k, colour = "red")
```
