---
title: "Supplement for Web Scraping 2"
author: "Mario Giacomazzo"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=F)
options(scipen=999)
library(tidyverse)    #Essential Functions
library(rvest)        #Read Tables From Webpages
```

# Introduction

This tutorial is the second part in our web scraping adventure. The R package `rvest` was developed by Hadley Wickham to provide a user-friendly web scraping experience. Combining `rvest` with a CSS selector tool, such as [SelectorGadget](https://selectorgadget.com/), or built-in browser developer tools, makes this entire process more efficient and less painful. This tutorial like the last will be guided and completed in stages during class. Remember the following lists linking to additional information on this topic.

* Blog Tutorials for Web Scraping in R Using `library(rvest)`.
    + **[Blog 1](http://blog.corynissen.com/2015/01/using-rvest-to-scrape-html-table.html)**
    + **[Blog 2](https://blog.rstudio.com/2014/11/24/rvest-easy-web-scraping-with-r/)**
    + **[Blog 3](http://categitau.com/using-rvest-to-scrape-data-from-wikipedia/)**
    + **[Blog 4](https://stat4701.github.io/edav/2015/04/02/rvest_tutorial/)**
    + **[Blog 5](https://www.analyticsvidhya.com/blog/2017/03/beginners-guide-on-web-scraping-in-r-using-rvest-with-hands-on-knowledge/)**
    
* Video Tutorials for Web Scraping in R Using `library(rvest)`
    +  **[Video 1](https://www.youtube.com/watch?v=4IYfYx4yoAI)**
    +  **[Video 2](https://www.youtube.com/watch?v=9ATgpE0yTxA)**
    +  **[Video 3](https://www.youtube.com/watch?v=gSbuwYdNYLM)**

* Information on Open Source SelectorGadget
    + **[Blog](https://cran.r-project.org/web/packages/rvest/vignettes/selectorgadget.html)**
    + **[Video](https://www.youtube.com/watch?v=A5LIHrpZBnY)**

In the last tutorial, we created 3 CSV files with primary and foreign keys. These files should be in the same folder as this RMarkdown file. The code below loads previously created CSV files into the global environment
```{r, message=F}
VIOLENT=read_csv("FINAL_VIOLENT.CSV")
ZIP=read_csv("FINAL_ZIP.CSV")
ABBREV=read_csv("FINAL_STATE_ABBREV.CSV")
```

The data frame `VIOLENT` links city and state names to violent crime statistics, the data frame `ZIP` links city and state abbreviations to latitude and longitude, and the data frame `ABBREV` links state names to state abbreviations. All 3 of these tables are required to join geographical locations to the violent crime statistics. 

# Part 1: Connection to Population Change and Density

Recall the Wikipedia page listing the [crime rates](https://en.wikipedia.org/wiki/List_of_United_States_cities_by_crime_rate) of US cities. This information comes from an FBI report from 2019. Suppose we wanted to investigate the effect of crime in 2019 on current population data in 2022. An HTML table, provided online, reports the [2022 population density and percentage change](http://worldpopulationreview.com/us-cities/) of 100 major US cities. Why is this information useful when related to violent crimes and what would you expect to discover?

## Chunk 1: Web Scraping and Variable Selection

We start by web scraping the population statistics table into R using functions from `rvest`. Additionally, we only select the variables of certain variables of interest. 

```{r,eval=F}
library(xml2)
URL.US_CENSUS = "http://worldpopulationreview.com/us-cities/"
US_CENSUS = URL.US_CENSUS %>%
              read_html() %>%
              html_table(header=T,fill=T) %>%
              .[[1]] %>%
              select(c(2:7) )

str(US_CENSUS)
```

## Chunk 2: Population Change and Density Variables to Numeric

Ideally, variables `Change` and `Density (mi²)` should be numeric. The following code starts by giving the variable `Density (mi²)` a more succinct name and then continues to provide the necessary mutations required to converting character vectors to numeric vectors. 

```{r,eval=F}
US_CENSUS2 = US_CENSUS %>%
                rename(Density=names(US_CENSUS)[6]) %>%
                mutate(Change=as.numeric(str_replace(Change,"%",""))) %>%
                mutate(Density=as.numeric(str_replace_all(Density,"[,]",""))) %>%
                mutate(`2022 Pop.`=as.numeric(str_replace_all(`2022 Pop.`,"[,]",""))) %>%
                mutate(`2010 Census`=as.numeric(str_replace_all(`2010 Census`,"[,]",""))) %>%
                mutate(Change=(`2022 Pop.`-`2010 Census`)/`2010 Census`*100)

write_csv(US_CENSUS2,"FINAL_CENSUS.CSV")
```


# Part 2: Inclusion of Expert Opinion

[SelectorGadget](https://selectorgadget.com/) is a point-and-click CSS selector useful for identifying the [CSS](https://en.wikipedia.org/wiki/Cascading_Style_Sheets)  or [XPath](https://en.wikipedia.org/wiki/XPath) to the specific portion of the webpage you want to scrape.

An article at https://www.securitysales.com/fire-intrusion/2018-safest-most-dangerous-states-us/ summarizes a study performed by [WalletHub](https://wallethub.com/edu/safest-states-to-live-in/4566/) which compared all 50 states using 48 key safety indicators. The article provides a link to the source along with lists providing the 10 most safest and 10 most dangerous states. 

## Chunk 1: Web Scraping Top 10 Most Safe and Least Safe States

Using [SelectorGadget](https://selectorgadget.com/), we point-and-click the text information we want on the webpage. This is the list of state names which comprise the most safe and most dangerous states. After deselecting the information we don't want, we can identify the CSS selector that links to the HTML element we want to scrape. Specifically for this case, the CSS selector is *#articleContentWrapper li*. The `html_nodes()` function in `rvest` locates in the HTML code where this CSS selector is being used to present the information visually on the webpage. Since this information is text, we use `html_text()` to read the list of states from the subsetted HTML code into a vector names `SAFE_VS_DANGEROUS`. 

```{r,eval=F}
URL.SAFE_VS_DANGEROUS = "https://www.securitysales.com/fire-intrusion/2018-safest-most-dangerous-states-us/"
SAFE_VS_DANGEROUS = URL.SAFE_VS_DANGEROUS %>%
                      read_html() %>%
                      html_nodes(css="#articleContentWrapper li") %>%
                      html_text() 


print(SAFE_VS_DANGEROUS)
```

## Chunk 2: Create a Data Frame Useful for Joining

Upon inspection of `SAFE_VS_DANGEROUS`, we notice that this vector of length *20* contains the top 10 most safe states and most dangerous states in the order they appear on the website. We create  new data frame combining these states with their classification, according to the website, that indicates whether a state is considered safe or dangerous. By linking this information to the violent crimes dataset, we can determine analytically the extent to which violent crimes were considered in this expert opinion. This may lead to either disagreement or agreement with the expert opinion. **The most important thing here is that we don't behave like the miserable bloggers who wave their flag of ignorance with anecdotal evidence.**

```{r,eval=F}
CLASS=c(rep("S", 10),rep("D", 10)) #S=SAFE and D=DANGEROUS
print(CLASS)
SAFE_VS_DANGEROUS2=as.tibble(cbind(SAFE_VS_DANGEROUS,CLASS))
names(SAFE_VS_DANGEROUS2)=c("STATE","CLASS")
write_csv(SAFE_VS_DANGEROUS2,"FINAL_SAFE_VS_DANGEROUS.CSV")
head(SAFE_VS_DANGEROUS2)
tail(SAFE_VS_DANGEROUS2)
```

